---
title: "PCGLM - JSS"
output:
  html_document:
    highlight: tango
    theme: simplex
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
vignette: |
  %\VignetteIndexEntry{my-vignette} %\VignetteEngine{knitr::rmarkdown}   %\VignetteEncoding{UTF-8}
---
  
  
```{r include=FALSE}
knitr::opts_chunk$set(
message=FALSE, warning=FALSE, paged.print=TRUE, collapse = TRUE, cache = TRUE, comment = "#>"
)
```


```{r setup}
library(pack)
library(mlogit)
devtools::load_all(".")
dist_ref <- new(ReferenceF)
dist_adj <- new(AdjacentR)
dist_cum <- new(CumulativeR)
dist_seq <- new(SequentialR)
```

Model Choice data just with the selected travel option

```{r}
{
  library(mlogit)
  data(ModeChoice, package = "Ecdat")
  head(ModeChoice)
  travel.long <- mlogit.data(ModeChoice, choice = "mode", shape = "long", alt.levels = c("air", "train", "bus", "car"))
  head(travel.long)
  choice <- sub(".*\\.", "", rownames(travel.long))
  indv <- sub("\\..*", "", rownames(travel.long))
  travel.long88 <- as.data.frame(cbind(indv, choice, travel.long))
}
travel_dat1 <- travel.long88[travel.long88$mode == T,]
head(travel_dat1)
```

### 3. Reference models for nominal response

REFERENCE, LOGISTIC, COMPLETE

```{r}
(l_1 <- dist_ref$GLMref(
  response = "choice",
  explanatory_complete = c("intercept", "hinc", "psize"),
  explanatory_proportional = c("NA"),
  distribution = "logistic",
  categories_order = c("air", "train", "bus", "car"),
  dataframe = travel_dat1
))
```

REFERENCE, LOGISTIC, PROPORTIONAL

```{r}
(l_2 <- dist_ref$GLMref(
  response = "choice",
  explanatory_complete = c("NA"),
  explanatory_proportional = c("intercept", "hinc", "psize"),
  distribution = "logistic",
  categories_order = c("air", "bus", "car", "train"),
  dataframe = travel_dat1
))
```

REFERENCE, CAUCHIT, COMPLETE
 
```{r}
(l_3 <- dist_ref$GLMref(
  response = "choice",
  explanatory_complete = c("intercept", "hinc", "psize"),
  explanatory_proportional = c("NA"),
  distribution = "cauchit",
  categories_order = c("air", "train", "bus", "car"),
  dataframe = travel_dat1
))
```

Then we change the reference category (bus as reference) and estimate again the three reference models:

REFERENCE, LOGISTIC, COMPLETE

```{r}
(l_1prime <- dist_ref$GLMref(
  response = "choice",
  explanatory_complete = c("intercept", "hinc", "psize"),
  explanatory_proportional = c("NA"),
  distribution = "logistic",
  categories_order = c("air", "train", "car", "bus"),
  dataframe = travel_dat1
))
```

REFERENCE, LOGISTIC, PROPORTIONAL

```{r}
(l_2prime <- dist_ref$GLMref(
  response = "choice",
  explanatory_complete = c("NA"),
  explanatory_proportional = c("intercept", "hinc", "psize"),
  distribution = "logistic",
  categories_order = c("air", "train","car", "bus"),
  dataframe = travel_dat1
))
```

REFERENCE, CAUCHIT, COMPLETE
 
```{r}
(l_3prime <- dist_ref$GLMref(
  response = "choice",
  explanatory_complete = c("intercept", "hinc", "psize"),
  explanatory_proportional = c("NA"),
  distribution = "cauchit",
  categories_order = c("air", "train","car", "bus"),
  dataframe = travel_dat1
))
```

The log-likelihoods l_1 and l_1prime are equal (since the canonical model is invariant under all permutations) whereas the log-likelihoods l_2 and l_2prime are different (respectively l_3 and l_3prime).

### 4. Adjacent models for ordinal response

Equivalence between (adjacent, logistic, complete) and (reference, logistic, complete) models.

REFERENCE, LOGISTIC, COMPLETE

```{r}
(estimation <- dist_ref$GLMref(
  response = "choice",
  explanatory_complete = c("intercept", "hinc", "psize"),
  explanatory_proportional = c("NA"),
  distribution = "logistic",
  categories_order = c("air", "bus", "car", "train"),
  dataframe = travel_dat1
))
```

ADJACENT, LOGISTIC, COMPLETE
 
```{r}
dist2 <- new(AdjacentR)
(estimation_prime <- dist_adj$GLMadj(
  response = "choice",
  explanatory_complete = c("intercept", "hinc", "psize"),
  explanatory_proportional = c("NA"),
  distribution = "logistic",
  categories_order = c("air", "bus", "car", "train"),
  dataframe = travel_dat1
))
```
Remark that the log-likelihoods l and l_prime are equal but the parameters estimations are different

#### Invariance under permutations

ADJACENT, CAUCHY, COMPLETE

```{r}
(estimation_1 <- dist_adj$GLMadj(
  response = "choice",
  explanatory_complete = c("intercept", "hinc", "psize"),
  explanatory_proportional = c("NA"),
  distribution = "cauchit",
  categories_order = c("air", "bus", "car", "train"),
  dataframe = travel_dat1
))
```

ADJACENT, GOMPERTZ, PROPORTIONAL
 
```{r}
(estimation_2 <- dist_adj$GLMadj(
  response = "choice",
  explanatory_complete = c("NA"),
  explanatory_proportional = c("intercept", "hinc", "psize"),
  distribution = "gompertz",
  categories_order = c("air", "bus", "car", "train"),
  dataframe = travel_dat1
))
```

ADJACENT, CAUCHY, COMPLETE (Reverse order)
 
```{r}
(estimation_1r <- dist_adj$GLMadj(
  response = "choice",
  explanatory_complete = c("intercept", "hinc", "psize"),
  explanatory_proportional = c("NA"),
  distribution = "cauchit",
  categories_order = c("train", "car", "bus", "air"),
  dataframe = travel_dat1
))
```

ADJACENT, GOMPERTZ, PROPORTIONAL (Reverse order)
 
```{r}
(estimation_2r <- dist_adj$GLMadj(
  response = "choice",
  explanatory_complete = c("NA"),
  explanatory_proportional = c("intercept", "hinc", "psize"),
  distribution = "gompertz",
  categories_order = c("train", "car", "bus", "air"),
  dataframe = travel_dat1
))
```

ADJACENT, GUMBEL, PROPORTIONAL (Reverse order)
 
```{r}
(estimation_3r <- dist_adj$GLMadj(
  response = "choice",
  explanatory_complete = c("NA"),
  explanatory_proportional = c("intercept", "hinc", "psize"),
  distribution = "gumbel",
  categories_order = c("train", "car", "bus", "air"),
  dataframe = travel_dat1
))
```

Remark that the log-likelihoods l_1 and l_1r are equal since the Cauchy distribution is symmetric whereas the log-likelihoods l_2 and l_2r are different since the Gompertz distribution is not symmetric. Moreover if the Gumbel distribution is used we the reverse order then the log-likelihoods l_2 and l_3r are equal since the Gumbel distribution is the symmetric of the Gompertz distribution. Otherwise, the parameter estimations are reversed.

### 5. Cumulative models for ordinal response

The equivalence between the (cumulative, Gompertz, proportional) and (sequential, Gompertz, proportional) models has been demonstrated by Läärä and Matthews (1985).

SEQUENTIAL, GOMPERTZ, PROPORTIONAL
 
```{r}
(estimation_prime <- dist_seq$GLMseq(
  response = "choice",
  explanatory_complete = c("NA"),
  explanatory_proportional = c("intercept", "hinc", "psize"),
  distribution = "gompertz",
  categories_order = c("air", "bus", "car", "train"),
  dataframe = travel_dat1
))
```

CUMULATIVE, GOMPERTZ, PROPORTIONAL (Error)
 
```{r}
# (estimation <- dist_cum$GLMcum(
#   response = "choice",
#   explanatory_complete = c("NA"),
#   explanatory_proportional = c("intercept", "hinc", "psize"),
#   distribution = "gompertz",
#   categories_order = c("air", "bus", "car", "train"),
#   dataframe = travel_dat1,   
#   beta_t = c("FALSE"),   beta_init = c(-0.08, 0.95, -0.84, -0.46, -0.18) 
# ))
# dat_lok_1 <- data.frame("LogLikIter" = estimation$LogLikIter, "iteration" = 1:length(estimation$LogLikIter))
# library(ggplot2)
# ggplot(dat_lok_1[-1,], aes(x = iteration, y = LogLikIter)) +
#   geom_point()
```


























