---
title: "Cumulative - R PCGLM"
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: simplex
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---


```{r}
dat <- read.csv2("~/Desktop/Test package/dat_cum.csv")
# summary(dat)
# str(dat)
head(dat)
devtools::load_all(".")
dist_cum <- new(CumulativeR)
```


### CUMULATIVE, LOGISTIC, COMPLETE (INTERCEPT)
```{r}
(estimation_prime <- dist_cum$GLMcum(
  response = "apply",
  explanatory_complete = c("intercept"),
  explanatory_proportional = c("NA"),
  distribution = "logistic",
  categories_order = c("unlikely", "somewhat likely", "very likely"),
  dataframe = dat,   
  beta_t = c("F"),
  beta_init = c(0.72, 2.76, 4.19, 5.06, -1.6, -0.85)
))
```

### CUMULATIVE, GOMPERTZ, COMPLETE (INTERCEPT, PARED)

```{r}
(estimation_prime <- dist_cum$GLMcum(
  response = "apply",
  explanatory_complete = c("intercept", "pared"),
  explanatory_proportional = c("NA"),
  distribution = "gompertz",
  categories_order = c("unlikely", "somewhat likely", "very likely"),
  dataframe = dat,
  beta_t = c("F"),   beta_init = c(0.72, 2.76, 4.19, 5.06, -1.6, -0.85) ))
```

### CUMULATIVE, GOMPERTZ, COMPLETE (INTERCEPT, PARED), PROPORTIONAL (PUBLIC)
DOES NOT WORK


```{r}
(estimation_prime <- dist_cum$GLMcum(
  response = "apply",
  explanatory_complete = c("intercept", "pared"),
  explanatory_proportional = c("public"),
  distribution = "gompertz",
  categories_order = c("unlikely", "somewhat likely", "very likely"),
  dataframe = dat,   
  beta_t = c("TRUE"),   beta_init = c(-0.08, 0.95, -0.84, -0.46, -0.18) ))
```

### CUMULATIVE, NORMAL, COMPLETE (INTERCEPT, PARED), PROPORTIONAL (PUBLIC)

```{r}
(estimation_prime <- dist_cum$GLMcum(
  response = "apply",
  explanatory_complete = c("intercept", "pared"),
  explanatory_proportional = c("public"),
  distribution = "normal",
  categories_order = c("unlikely", "somewhat likely", "very likely"),
  dataframe = dat,   
  beta_t = c("F"),   beta_init = c(0.72, 2.76, 4.19, 5.06, -1.6, -0.85) ))
```

### CUMULATIVE, NORMAL, COMPLETE (PARED, GPA), PROPORTIONAL (INTERCEPT, PUBLIC)

```{r}
(estimation_prime <- dist_cum$GLMcum(
  response = "apply",
  explanatory_complete = c("pared", "gpa"),
  explanatory_proportional = c("intercept","public"),
  distribution = "normal",
  categories_order = c("unlikely", "somewhat likely", "very likely"),
  dataframe = dat,
  beta_t = c("F"),   beta_init = c(0.72, 2.76, 4.19, 5.06, -1.6, -0.85) ))
```

### CUMULATIVE, NORMAL, COMPLETE (INTERCEPT, PARED), PROPORTIONAL (PUBLIC, GPA)
DOES NOT WORK
```{r error=TRUE}
(estimation_prime <- dist_cum$GLMcum(
  response = "apply",
  explanatory_complete = c("intercept", "pared"),
  explanatory_proportional = c("public", "gpa"),
  distribution = "normal",
  categories_order = c("unlikely", "somewhat likely", "very likely"),
  dataframe = dat,   beta_t = c("F"),   beta_init = c(0.72, 2.76, 4.19, 5.06, -1.6, -0.85) ))
```

### CUMULATIVE, CAUCHIT, COMPLETE (INTERCEPT, PARED), PROPORTIONAL (PUBLIC)

DOES NOT WORK 

```{r}
(estimation_prime <- dist_cum$GLMcum(
  response = "apply",
  explanatory_complete = c("intercept", "pared"),
  explanatory_proportional = c("public"),
  distribution = "cauchit",
  categories_order = c("unlikely", "somewhat likely", "very likely"),
  dataframe = dat,   beta_t = c("F"),   beta_init = c(0.72, 2.76, 4.19, 5.06, -1.6, -0.85) ))
```


